---
title: "Homework 6"
author: "Vihar Desu"
date: '`r format(Sys.time(), "%Y-%m-%d")`'
output: github_document
---

## Setup 

##### General
```{r general_setup}
knitr::opts_chunk$set(
  echo = TRUE,
  include = TRUE,
  message = FALSE,
  warning = FALSE
)
```

##### Installations
```{r installations}
library(tidyverse)
library(ggridges)
library(patchwork)
library(readxl)
library(viridis)
library(modelr)
library(p8105.datasets)
```

##### Visualizations
```{r visualiztion_setup}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

### Problem 1

```{r}
homicide_df = 
  read_csv("data/homicide-data.csv", na = c("", "NA", "Unknown")) %>% 
  mutate(
    city_state = str_c(city, state, sep = ", "),
    victim_age = as.numeric(victim_age),
    resolution = case_when(
      disposition == "Closed without arrest" ~ 0,
      disposition == "Open/No arrest"        ~ 0,
      disposition == "Closed by arrest"      ~ 1)
  ) %>% 
  filter(
    victim_race %in% c("White", "Black"),
    !city_state %in% c("Tulsa, AL", "Dallas, TX","Phoenix, AZ", "Kansas City, MO")) %>% 
  select(city_state, resolution, victim_age, victim_race, victim_sex)
```

```{r}
knitr::kable(head(homicide_df), "simple", caption="Source: Homicide Dataset")
```

```{r}
baltimore_df =
  homicide_df %>% 
  filter(city_state == "Baltimore, MD")

glm(resolution ~ victim_age + victim_race + victim_sex, 
    data = baltimore_df,
    family = binomial()) %>% 
  broom::tidy() %>% 
  mutate(
    OR = exp(estimate),
    CI_lower = exp(estimate - 1.96 * std.error),
    CI_upper = exp(estimate + 1.96 * std.error)
  ) %>% 
  select(term, OR, starts_with("CI")) %>% 
  knitr::kable(digits = 3, caption="Source: Homicide Dataset")
```

```{r}
models_results_df = 
  homicide_df %>% 
  nest(data = -city_state) %>% 
  mutate(
    models = 
      map(.x = data, ~glm(resolution ~ victim_age + victim_race + victim_sex, data = .x, family = binomial())),
    results = map(models, broom::tidy)
  ) %>% 
  select(city_state, results) %>% 
  unnest(results) %>% 
  mutate(
    OR = exp(estimate),
    CI_lower = exp(estimate - 1.96 * std.error),
    CI_upper = exp(estimate + 1.96 * std.error)
  ) %>% 
  select(city_state, term, OR, starts_with("CI")) 
```

```{r}
models_results_df %>% 
  filter(term == "victim_sexMale") %>% 
  mutate(city_state = fct_reorder(city_state, OR)) %>% 
  ggplot(aes(x = city_state, y = OR)) + 
  geom_point() + 
  geom_errorbar(aes(ymin = CI_lower, ymax = CI_upper)) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

The results read as follows:  
The odds that a homicide goes unresolved if the victim is black is X times greater (or less than) that of the same odds if the victim were white. The results are stratified by city. From our analysis, we see mostly tight confidence intervals for our adjusted odds ratios. Albuquerque, New Mexico has the highest confidence interval and odds ratio while New York, NY has the lowest.  
---

## Problem 2

```{r}
birthweight_df = 
  read_csv("data/birthweight.csv", na = c("", "NA", "Unknown")) %>% 
  mutate(
    babysex = case_when(
      babysex == 1 ~ "male",
      babysex == 2 ~ "female"
    ),
    malform = case_when(
      malform == 0 ~ "absent",
      malform == 1 ~ "present"
    ),
    frace = case_when(
      frace == 1 ~ "white",
      frace == 2 ~ "black",
      frace == 3 ~ "asian",
      frace == 4 ~ "puerto_rican",
      frace == 8 ~ "other",
      frace == 9 ~ "unknown"
    ),
    mrace = case_when(
      mrace == 1 ~ "white",
      mrace == 2 ~ "black",
      mrace == 3 ~ "asian",
      mrace == 4 ~ "puerto_rican",
      mrace == 8 ~ "other",
      mrace == 9 ~ "unknown"
    )
  )

birthweight_df
```



